---
title: "Using Your Own Data"
author: "Olatunde D. Akanbi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Your Own Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

```{r setup}
library(manureshed)
```

## Overview

The `manureshed` package includes data from 1987-2016, but you can add your own:

- **WWTP data** for years outside 2007-2016
- **Agricultural data** from other sources  
- **Custom spatial boundaries**
- **Industrial or urban nutrient sources**

## Using Custom WWTP Data

### For Years Outside 2007-2016

The package has WWTP data for 2007-2016. For other years, provide your own:

```{r custom_wwtp_basic, eval=FALSE}
# Use your WWTP files for 2020
results_2020 <- run_builtin_analysis(
  scale = "huc8",
  year = 2020,  # Agricultural data available 1987-2016
  nutrients = "nitrogen",
  include_wwtp = TRUE,
  custom_wwtp_nitrogen = "nitrogen_2021.csv",
  wwtp_load_units = "kg"
)

# 3. Check results
summarize_results(results_custom)
quick_check(results_custom)

# 4. Create maps
nitrogen_map <- map_agricultural_classification(
  results_custom$integrated$nitrogen,
  "nitrogen", "combined_N_class",
  "2021 Nitrogen with Custom WWTP Data"
)

save_plot(nitrogen_map, "custom_analysis_2021.png")

# 5. Save results
save_spatial_data(
  results_custom$integrated$nitrogen,
  "nitrogen_2021_results.rds"
)
```

This example shows the complete workflow from custom data to final maps and saved results.

## Advanced Custom Data Integration

### Adding Other Nutrient Sources

You can integrate additional nutrient sources beyond WWTP:

```{r other_sources, eval=FALSE}
# Example: Adding industrial sources
industrial_data <- data.frame(
  Facility_Name = c("Steel Plant A", "Chemical Plant B", "Food Processor C"),
  Lat = c(41.5, 40.7, 42.1),
  Long = c(-81.7, -82.1, -83.5),
  N_Load_tons = c(50, 75, 30),
  P_Load_tons = c(10, 15, 8),
  Source_Type = "Industrial"
)

# Convert to spatial format
industrial_sf <- st_as_sf(industrial_data, 
                         coords = c("Long", "Lat"), 
                         crs = 4326) %>%
  st_transform(5070)  # Transform to analysis CRS

# Aggregate by spatial units (example for HUC8)
boundaries <- load_builtin_boundaries("huc8")
industrial_aggregated <- wwtp_aggregate_by_boundaries(
  industrial_sf, boundaries, "nitrogen", "huc8"
)

# Add to existing results
# (You would modify the integration functions to include industrial sources)
```

### Working with Different Time Periods

```{r time_periods, eval=FALSE}
# Create a time series dataset
years_to_analyze <- 2018:2022
time_series_results <- list()

for (year in years_to_analyze) {
  # Check if you have WWTP data for this year
  wwtp_file <- paste0("wwtp_nitrogen_", year, ".csv")
  
  if (file.exists(wwtp_file)) {
    time_series_results[[as.character(year)]] <- run_builtin_analysis(
      scale = "huc8",
      year = year,
      nutrients = "nitrogen",
      include_wwtp = TRUE,
      custom_wwtp_nitrogen = wwtp_file,
      save_outputs = FALSE,
      verbose = FALSE
    )
  } else {
    # Agricultural only if no WWTP data
    time_series_results[[as.character(year)]] <- run_builtin_analysis(
      scale = "huc8", 
      year = year,
      nutrients = "nitrogen",
      include_wwtp = FALSE,
      save_outputs = FALSE,
      verbose = FALSE
    )
  }
}

# Compare results across years
yearly_summary <- do.call(rbind, lapply(names(time_series_results), function(year) {
  result <- time_series_results[[year]]
  classes <- table(result$agricultural$N_class)
  
  data.frame(
    Year = as.numeric(year),
    Total_Units = sum(classes),
    Source_Units = classes["Source"] %||% 0,
    Sink_Units = sum(classes[c("Sink_Deficit", "Sink_Fertilizer")], na.rm = TRUE),
    Excluded_Units = classes["Excluded"] %||% 0
  )
}))

print(yearly_summary)
```

### Custom Agricultural Data

If you have agricultural data from other sources:

```{r custom_agricultural, eval=FALSE}
# Example: Custom agricultural data format
custom_farm_data <- data.frame(
  County_FIPS = c("39001", "39003", "39005"),
  Manure_N_kg = c(100000, 150000, 200000),
  Manure_P_kg = c(20000, 30000, 40000),
  Fertilizer_N_kg = c(50000, 75000, 100000),
  Fertilizer_P_kg = c(10000, 15000, 20000),
  N_Fixation_kg = c(25000, 40000, 35000),
  Crop_N_Removal_kg = c(80000, 120000, 140000),
  Crop_P_Removal_kg = c(15000, 25000, 30000),
  Cropland_acres = c(45000, 67000, 52000)
)

# Convert to package format
standardized_farm_data <- data.frame(
  ID = custom_farm_data$County_FIPS,
  NAME = paste("County", substr(custom_farm_data$County_FIPS, 3, 5)),
  manure_N = custom_farm_data$Manure_N_kg,
  manure_P = custom_farm_data$Manure_P_kg,
  fertilizer_N = custom_farm_data$Fertilizer_N_kg,
  fertilizer_P = custom_farm_data$Fertilizer_P_kg,
  N_fixation = custom_farm_data$N_Fixation_kg,
  crop_removal_N = custom_farm_data$Crop_N_Removal_kg,
  crop_removal_P = custom_farm_data$Crop_P_Removal_kg,
  cropland = custom_farm_data$Cropland_acres
)

# Apply classifications
custom_classified <- standardized_farm_data %>%
  agri_classify_nitrogen(cropland_threshold = 1234, scale = "county") %>%
  agri_classify_phosphorus(cropland_threshold = 1234, scale = "county")

print(custom_classified)
```

### Data Validation and Quality Control

```{r data_validation, eval=FALSE}
# Function to validate your custom data
validate_custom_data <- function(data, data_type = "wwtp") {
  
  issues <- list()
  
  if (data_type == "wwtp") {
    # Check required columns
    required_cols <- c("Facility_Name", "Lat", "Long")
    missing_cols <- setdiff(required_cols, names(data))
    if (length(missing_cols) > 0) {
      issues$missing_columns <- missing_cols
    }
    
    # Check coordinate ranges (for US)
    if ("Lat" %in% names(data) && "Long" %in% names(data)) {
      invalid_coords <- sum(data$Lat < 24 | data$Lat > 50 | 
                           data$Long < -125 | data$Long > -66, na.rm = TRUE)
      if (invalid_coords > 0) {
        issues$invalid_coordinates <- invalid_coords
      }
    }
    
    # Check for negative loads
    load_cols <- names(data)[grepl("Load|load", names(data))]
    for (col in load_cols) {
      if (col %in% names(data)) {
        negative_loads <- sum(data[[col]] < 0, na.rm = TRUE)
        if (negative_loads > 0) {
          issues[[paste0("negative_", col)]] <- negative_loads
        }
      }
    }
  }
  
  if (data_type == "agricultural") {
    # Check for negative values in nutrient data
    nutrient_cols <- c("manure_N", "manure_P", "fertilizer_N", "fertilizer_P",
                       "crop_removal_N", "crop_removal_P", "cropland")
    
    for (col in nutrient_cols) {
      if (col %in% names(data)) {
        negative_values <- sum(data[[col]] < 0, na.rm = TRUE)
        if (negative_values > 0) {
          issues[[paste0("negative_", col)]] <- negative_values
        }
      }
    }
  }
  
  if (length(issues) == 0) {
    message("✓ Data validation passed")
  } else {
    message("⚠ Data validation issues found:")
    for (issue in names(issues)) {
      message("  ", issue, ": ", issues[[issue]])
    }
  }
  
  return(issues)
}

# Use the validation function
# validate_custom_data(your_wwtp_data, "wwtp")
# validate_custom_data(your_farm_data, "agricultural")
```

### Exporting Results

```{r export_results, eval=FALSE}
# Export results in different formats
export_analysis_results <- function(results, output_dir = "exports") {
  
  dir.create(output_dir, showWarnings = FALSE)
  
  # Export agricultural data as CSV
  agri_csv <- st_drop_geometry(results$agricultural)
  write.csv(agri_csv, file.path(output_dir, "agricultural_results.csv"), row.names = FALSE)
  
  # Export as shapefile (if sf package available)
  if (requireNamespace("sf", quietly = TRUE)) {
    st_write(results$agricultural, file.path(output_dir, "agricultural_results.shp"), 
             delete_dsn = TRUE, quiet = TRUE)
  }
  
  # Export WWTP facilities if available
  if ("wwtp" %in% names(results)) {
    for (nutrient in names(results$wwtp)) {
      facility_data <- results$wwtp[[nutrient]]$facility_data
      write.csv(facility_data, 
                file.path(output_dir, paste0("wwtp_", nutrient, "_facilities.csv")), 
                row.names = FALSE)
    }
  }
  
  # Export integrated results if available
  if ("integrated" %in% names(results)) {
    for (nutrient in names(results$integrated)) {
      integrated_csv <- st_drop_geometry(results$integrated[[nutrient]])
      write.csv(integrated_csv, 
                file.path(output_dir, paste0("integrated_", nutrient, "_results.csv")), 
                row.names = FALSE)
    }
  }
  
  message("Results exported to: ", output_dir)
}

# Use the export function
# export_analysis_results(results_custom, "my_exports")
```

## Troubleshooting Common Issues

### WWTP Data Issues

```{r troubleshooting_wwtp, eval=FALSE}
# Common WWTP data problems and solutions

# Problem 1: "No valid facilities remaining after cleaning"
# Solution: Check coordinates and load values
wwtp_data <- read.csv("your_wwtp_file.csv")
summary(wwtp_data)  # Look for obvious issues

# Check coordinate ranges
summary(wwtp_data$Latitude)   # Should be ~24-50 for US
summary(wwtp_data$Longitude)  # Should be ~-125 to -66 for US

# Check load values
summary(wwtp_data$Load)  # Should be positive numbers

# Problem 2: Wrong coordinate system
# Solution: Make sure coordinates are in decimal degrees (not UTM)
# Example conversion if needed:
# library(sf)
# points_utm <- st_as_sf(data, coords = c("X_UTM", "Y_UTM"), crs = 32617)
# points_dd <- st_transform(points_utm, 4326)
# coords_dd <- st_coordinates(points_dd)

# Problem 3: Mixed units in same file
# Solution: Standardize units before analysis
standardize_mixed_units <- function(data, load_col, unit_col) {
  data$standardized_load <- ifelse(
    data[[unit_col]] == "kg", data[[load_col]],
    ifelse(data[[unit_col]] == "lbs", data[[load_col]] / 2.20462,
           data[[load_col]] * 907.185)  # assuming tons
  )
  return(data)
}
```

### Agricultural Data Issues

```{r troubleshooting_agricultural, eval=FALSE}
# Common agricultural data problems

# Problem: Impossible nutrient balances
# Solution: Check your units and conversion factors
check_nutrient_balance <- function(data) {
  # Calculate simple checks
  data$total_inputs_N <- data$manure_N + data$fertilizer_N + data$N_fixation
  data$efficiency_N <- data$crop_removal_N / data$total_inputs_N
  
  # Flag suspicious values
  suspicious <- data$efficiency_N > 2.0 | data$efficiency_N < 0.1
  
  if (sum(suspicious, na.rm = TRUE) > 0) {
    message("Warning: ", sum(suspicious, na.rm = TRUE), " units have suspicious N efficiency")
    print(data[suspicious, c("ID", "total_inputs_N", "crop_removal_N", "efficiency_N")])
  }
  
  return(data)
}

# Problem: Missing spatial boundaries
# Solution: Use built-in boundaries or provide your own
# county_boundaries <- load_builtin_boundaries("county")
# your_data_with_boundaries <- merge(your_data, county_boundaries, by.x = "FIPS", by.y = "FIPS")
```

## Getting Help

### Common Questions

**Q: What format should my WWTP data be in?**
A: CSV file with columns for facility name, latitude, longitude, annual load, and state. The package can auto-detect most EPA formats.

**Q: What units can I use for loads?**  
A: "kg", "lbs", "pounds", or "tons". Specify with `wwtp_load_units`.

**Q: Can I use data from other countries?**
A: The spatial boundaries are US-only, but you could provide your own boundaries and modify the coordinate system.

**Q: How do I handle missing data?**
A: The package excludes facilities with missing coordinates or loads. Clean your data first for best results.

**Q: My analysis is running slowly. What can I do?**
A: Try using a smaller spatial scale (HUC2 < HUC8 < County in terms of processing time), fewer years, or clear the data cache with `clear_data_cache()`.

**Q: How do I cite the package and data?**
A: Use `citation_info()` to get proper citations for the package and underlying datasets.

### Function Help

```{r help, eval=FALSE}
# Get help on specific functions
?load_user_wwtp
?run_builtin_analysis  
?wwtp_clean_data

# Check what data is available
check_builtin_data()
list_available_years()

# Package health check
health_check()

# Test data download connection
test_osf_connection()
```

### Getting More Help

- Check the other vignettes: `vignette("getting-started")`, `vignette("visualization-guide")`, `vignette("advanced-features")`
- Look at function documentation: `?function_name`
- Check the package website or GitHub repository for examples
- Use `quick_check()` to validate your results

The `manureshed` package is designed to work with your data as easily as possible. Start with the built-in examples, then gradually add your own data as needed.nitrogen_2020.csv"
)

### Handling Different Data Formats

#### Standard EPA Format

Most EPA WWTP files work automatically:

```{r epa_standard, eval=FALSE}
# Standard EPA format (auto-detected)
results <- run_builtin_analysis(
  scale = "county",
  year = 2018,
  nutrients = c("nitrogen", "phosphorus"),
  include_wwtp = TRUE,
  custom_wwtp_nitrogen = "nitrogen_2018.csv",
  custom_wwtp_phosphorus = "phosphorus_2018.csv"
)
```

#### Different Units

```{r different_units, eval=FALSE}
# If your data uses pounds instead of kg
results <- run_builtin_analysis(
  scale = "huc8", 
  year = 2019,
  nutrients = "nitrogen",
  include_wwtp = TRUE,
  custom_wwtp_nitrogen = "nitrogen_pounds_2019.csv",
  wwtp_load_units = "lbs"  # Converts automatically
)

# Other units: "kg", "lbs", "pounds", "tons"
```

#### Different File Format

```{r different_format, eval=FALSE}
# If headers are in different rows
results <- run_builtin_analysis(
  scale = "county",
  year = 2021, 
  nutrients = "nitrogen",
  include_wwtp = TRUE,
  custom_wwtp_nitrogen = "nitrogen_2021.csv",
  wwtp_skip_rows = 3,      # Skip first 3 rows
  wwtp_header_row = 1      # Headers in row 1 after skipping
)
```

#### Custom Column Names

```{r custom_columns, eval=FALSE}
# If your columns have different names
custom_mapping <- list(
  facility_name = "Plant_Name",
  latitude = "Lat_DD",
  longitude = "Long_DD", 
  pollutant_load = "Annual_Load_kg",
  state = "State_Code"
)

results <- run_builtin_analysis(
  scale = "huc8",
  year = 2022,
  nutrients = "nitrogen", 
  include_wwtp = TRUE,
  custom_wwtp_nitrogen = "custom_format.csv",
  wwtp_column_mapping = custom_mapping
)
```

### Manual WWTP Processing

For full control, process WWTP data step by step:

```{r manual_wwtp, eval=FALSE}
# Step 1: Load your WWTP file
wwtp_raw <- load_user_wwtp(
  file_path = "nitrogen_2020.csv",
  nutrient = "nitrogen",
  load_units = "lbs"
)

# Step 2: Clean the data
wwtp_clean <- wwtp_clean_data(wwtp_raw, "nitrogen")

# Step 3: Classify facilities by size
wwtp_classified <- wwtp_classify_sources(wwtp_clean, "nitrogen")

# Step 4: Convert to spatial format
wwtp_spatial <- wwtp_to_spatial(wwtp_classified)

# Step 5: Load boundaries and aggregate by spatial units
boundaries <- load_builtin_boundaries("huc8")
wwtp_aggregated <- wwtp_aggregate_by_boundaries(
  wwtp_spatial, boundaries, "nitrogen", "huc8"
)

# Step 6: Integrate with agricultural data
agri_data <- load_builtin_nugis("huc8", 2020)
agri_processed <- agri_classify_complete(agri_data, "huc8")

integrated <- integrate_wwtp_agricultural(
  agri_processed, wwtp_aggregated, "nitrogen", 
  cropland_threshold = 1234, scale = "huc8"
)
```

## Unit Conversions

### Common Conversions

```{r unit_conversions, eval=FALSE}
# Convert between units
kg_loads <- c(1000, 2000, 5000)
tons_loads <- convert_load_units(kg_loads, "kg")

lbs_loads <- c(2000, 4000, 10000)  
tons_loads <- convert_load_units(lbs_loads, "lbs")

print(data.frame(
  Original_kg = kg_loads,
  Converted_tons = convert_load_units(kg_loads, "kg")
))
```

### Handling P2O5 vs P

The package automatically converts P2O5 to P, but you can do it manually:

```{r p2o5_conversion, eval=FALSE}
# If you have P2O5 data, convert to P
p2o5_values <- c(100, 200, 300)  # kg P2O5
p_values <- p2o5_values * P2O5_TO_P  # Convert to P

print(paste("P2O5 to P conversion factor:", P2O5_TO_P))
```

## Working with Different Spatial Scales

### County Data (FIPS Codes)

```{r county_data, eval=FALSE}
# County analysis - make sure you have 5-digit FIPS codes
county_results <- run_builtin_analysis(
  scale = "county",
  year = 2016,
  nutrients = "nitrogen"
)

# Your county WWTP data should have state and county info
```

### HUC8 Watersheds

```{r huc8_data, eval=FALSE}
# HUC8 analysis - 8-digit watershed codes
huc8_results <- run_builtin_analysis(
  scale = "huc8", 
  year = 2016,
  nutrients = "nitrogen"
)

# Make sure HUC8 codes are 8 digits (add leading zero if needed)
huc8_codes <- c("4110001", "4110002")  # 7 digits
formatted_codes <- format_huc8(huc8_codes)  # Adds leading zero
print(formatted_codes)  # "04110001", "04110002"
```

### HUC2 Regions

```{r huc2_data, eval=FALSE}
# HUC2 analysis - 2-digit regional codes
huc2_results <- run_builtin_analysis(
  scale = "huc2",
  year = 2016, 
  nutrients = "nitrogen"
)
```

## State-Specific Analysis

### Single State

```{r state_analysis, eval=FALSE}
# Analyze just one state
iowa_results <- run_state_analysis(
  state = "IA",
  scale = "county",
  year = 2016,
  nutrients = "nitrogen",
  include_wwtp = TRUE
)

# With custom WWTP data
texas_results <- run_state_analysis(
  state = "TX",
  scale = "huc8",
  year = 2020,
  nutrients = "nitrogen", 
  include_wwtp = TRUE,
  custom_wwtp_nitrogen = "texas_wwtp_2020.csv"
)
```

### Multiple States

```{r multi_state, eval=FALSE}
# Analyze several states
midwest_states <- c("IA", "IL", "IN", "OH") 
state_results <- list()

for (state in midwest_states) {
  state_results[[state]] <- run_state_analysis(
    state = state,
    scale = "county",
    year = 2016,
    nutrients = "nitrogen", 
    include_wwtp = TRUE
  )
}
```

## Quality Control

### Validate Your Data

```{r data_validation2, eval=FALSE}
# Check your results make sense
quick_check(results)

# Look at the classifications
table(results$agricultural$N_class)

# Check WWTP facility counts
if ("wwtp" %in% names(results)) {
  print(paste("WWTP facilities:", nrow(results$wwtp$nitrogen$facility_data)))
}
```

### Common Data Issues

```{r data_issues, eval=FALSE}
# Problem: Negative nutrient values
# Solution: Check your data source and units

# Problem: All facilities excluded  
# Solution: Check coordinate system (should be decimal degrees)

# Problem: No WWTP facilities found
# Solution: Check facility coordinates are in correct format

# Problem: Classification doesn't make sense
# Solution: Check cropland threshold and nutrient balance calculations
```

## Working with Multiple Years

### Time Series Analysis

```{r time_series, eval=FALSE}
# Analyze multiple years
years_to_analyze <- 2014:2016

batch_results <- batch_analysis_years(
  years = years_to_analyze,
  scale = "huc8",
  nutrients = "nitrogen", 
  include_wwtp = TRUE
)

# With custom WWTP data for some years
custom_wwtp_files <- list(
  "2018" = "nitrogen_2018.csv",
  "2019" = "nitrogen_2019.csv", 
  "2020" = "nitrogen_2020.csv"
)

# Process each year with custom data
custom_results <- list()
for (year in names(custom_wwtp_files)) {
  custom_results[[year]] <- run_builtin_analysis(
    scale = "county",
    year = as.numeric(year),
    nutrients = "nitrogen",
    include_wwtp = TRUE,
    custom_wwtp_nitrogen = custom_wwtp_files[[year]]
  )
}
```

## Data Management Tips

### File Organization

```{r , eval=FALSE}
# Organize your data files
# project_folder/
#   ├── wwtp_data/
#   │   ├── nitrogen_2018.csv
#   │   ├── nitrogen_2019.csv
#   │   └── phosphorus_2018.csv
#   ├── results/
#   └── maps/

# Set working directory
setwd("project_folder")

# Run analysis with organized files
results <- run_builtin_analysis(
  scale = "huc8",
  year = 2018,
  nutrients = "nitrogen",
  include_wwtp = TRUE,
  custom_wwtp_nitrogen = "wwtp_data/nitrogen_2018.csv",
  output_dir = "results"
)
```

### Memory Management

```{r memory_management, eval=FALSE}
# For large datasets, clear cache periodically
clear_data_cache()

# Check package health
health_check()

# Free up memory between analyses
rm(large_results)
gc()
```

## Example: Complete Custom Workflow

Here's a complete example using custom data:

```{r complete_example, eval=FALSE}
# 1. Prepare your WWTP file (nitrogen_2021.csv)
# Make sure it has columns: Facility Name, Latitude, Longitude, Load (kg/yr), State

# 2. Run analysis
results_custom <- run_builtin_analysis(
  scale = "huc8",
  year = 2021,  # Agricultural data extrapolated to 2021
  nutrients = "nitrogen",
  include_wwtp = TRUE,
  custom_wwtp_nitrogen = "nitrogen_2021.csv",
  wwtp_load_units = "kg"
)

# 3. Check results
summarize_results(results_custom)
quick_check(results_custom)

# 4. Create maps
nitrogen_map <- map_agricultural_classification(
  results_custom$integrated$nitrogen,
  "nitrogen", "combined_N_class",
  "2021 Nitrogen with Custom WWTP Data"
)

save_plot(nitrogen_map, "custom_analysis_2021.png")

# 5. Save results
save_spatial_data(
  results_custom$integrated$nitrogen,
  "nitrogen_2021_results.rds"
)
```








### Integration Issues

```{r troubleshooting_integration, eval=FALSE}
# Common integration problems

# Problem: WWTP facilities not matching spatial units
# Solution: Check coordinate systems and boundary files
check_wwtp_spatial_match <- function(wwtp_data, boundaries) {
  
  # Convert WWTP to spatial
  wwtp_sf <- st_as_sf(wwtp_data, coords = c("Long", "Lat"), crs = 4326)
  wwtp_projected <- st_transform(wwtp_sf, st_crs(boundaries))
  
  # Check spatial intersection
  intersections <- st_intersects(wwtp_projected, boundaries)
  
  # Count facilities per spatial unit
  matches <- lengths(intersections)
  
  message("WWTP spatial matching summary:")
  message("  Total facilities: ", nrow(wwtp_data))
  message("  Facilities matched to boundaries: ", sum(matches > 0))
  message("  Facilities not matched: ", sum(matches == 0))
  
  if (sum(matches == 0) > 0) {
    unmatched <- wwtp_data[matches == 0, ]
    message("  Check coordinates for unmatched facilities")
    print(head(unmatched))
  }
  
  return(matches)
}

# Problem: Scale mismatch between data sources
# Solution: Ensure consistent spatial scales
verify_scale_consistency <- function(agricultural_data, wwtp_data, scale) {
  
  message("Scale consistency check for: ", scale)
  
  # Check ID formats
  if (scale == "county") {
    # FIPS codes should be 5 digits
    invalid_fips <- sum(nchar(agricultural_data$ID) != 5)
    message("  Invalid FIPS codes: ", invalid_fips)
  } else if (scale == "huc8") {
    # HUC8 codes should be 8 digits
    invalid_huc8 <- sum(nchar(agricultural_data$ID) != 8)
    message("  Invalid HUC8 codes: ", invalid_huc8)
  }
  
  # Check coordinate ranges
  coord_summary <- list(
    lat_range = range(wwtp_data$Lat, na.rm = TRUE),
    long_range = range(wwtp_data$Long, na.rm = TRUE)
  )
  
  message("  WWTP coordinate ranges:")
  message("    Latitude: ", paste(round(coord_summary$lat_range, 2), collapse = " to "))
  message("    Longitude: ", paste(round(coord_summary$long_range, 2), collapse = " to "))
  
  return(coord_summary)
}
```

## Best Practices for Custom Data

### File Organization

```{r file_organization, eval=FALSE}
# Recommended project structure
create_project_structure <- function(project_name) {
  
  # Create main directories
  dir.create(project_name, showWarnings = FALSE)
  dir.create(file.path(project_name, "data"), showWarnings = FALSE)
  dir.create(file.path(project_name, "data", "wwtp"), showWarnings = FALSE)
  dir.create(file.path(project_name, "data", "agricultural"), showWarnings = FALSE)
  dir.create(file.path(project_name, "results"), showWarnings = FALSE)
  dir.create(file.path(project_name, "maps"), showWarnings = FALSE)
  dir.create(file.path(project_name, "exports"), showWarnings = FALSE)
  
  # Create README
  readme_content <- paste(
    "# Manureshed Analysis Project",
    "",
    "## Data Files",
    "- data/wwtp/ - WWTP facility data files",
    "- data/agricultural/ - Custom agricultural data",
    "",
    "## Outputs", 
    "- results/ - Analysis results (.rds files)",
    "- maps/ - Generated maps (.png files)",
    "- exports/ - Exported data (.csv, .shp files)",
    "",
    "## Analysis Scripts",
    "- analysis.R - Main analysis script",
    "",
    sep = "\n"
  )
  
  writeLines(readme_content, file.path(project_name, "README.md"))
  
  message("Project structure created: ", project_name)
  message("Add your data files to the appropriate folders")
}

# Create organized project
# create_project_structure("my_nutrient_analysis")
```

### Data Documentation

```{r data_documentation, eval=FALSE}
# Document your custom data sources
document_data_sources <- function(wwtp_files = NULL, agricultural_files = NULL, 
                                 output_file = "data_documentation.txt") {
  
  doc_content <- c(
    "DATA SOURCES DOCUMENTATION",
    "============================",
    "",
    "Analysis Date: ", as.character(Sys.Date()),
    "Package Version: ", as.character(packageVersion("manureshed")),
    ""
  )
  
  if (!is.null(wwtp_files)) {
    doc_content <- c(doc_content,
      "WWTP DATA FILES:",
      "----------------"
    )
    
    for (file in wwtp_files) {
      if (file.exists(file)) {
        file_info <- file.info(file)
        doc_content <- c(doc_content,
          paste("File:", file),
          paste("  Size:", round(file_info$size / 1024, 2), "KB"),
          paste("  Modified:", file_info$mtime),
          paste("  Rows:", nrow(read.csv(file))),
          ""
        )
      }
    }
  }
  
  if (!is.null(agricultural_files)) {
    doc_content <- c(doc_content,
      "AGRICULTURAL DATA FILES:",
      "------------------------"
    )
    
    for (file in agricultural_files) {
      if (file.exists(file)) {
        file_info <- file.info(file)
        doc_content <- c(doc_content,
          paste("File:", file),
          paste("  Size:", round(file_info$size / 1024, 2), "KB"),
          paste("  Modified:", file_info$mtime),
          paste("  Rows:", nrow(read.csv(file))),
          ""
        )
      }
    }
  }
  
  doc_content <- c(doc_content,
    "ANALYSIS PARAMETERS:",
    "--------------------",
    "Built-in data years: 1987-2016 (Agricultural), 2007-2016 (WWTP)",
    "Spatial scales: county, huc8, huc2",
    "Default cropland threshold: 500 ha (1,234 acres)",
    ""
  )
  
  writeLines(doc_content, output_file)
  message("Data documentation saved to: ", output_file)
}

# Use documentation function
# document_data_sources(
#   wwtp_files = c("data/wwtp/nitrogen_2020.csv", "data/wwtp/phosphorus_2020.csv"),
#   agricultural_files = c("data/agricultural/custom_farm_data.csv")
# )
```

### Quality Assurance Workflow

```{r qa_workflow, eval=FALSE}
# Complete quality assurance workflow
quality_assurance_workflow <- function(results, data_sources = NULL) {
  
  qa_report <- list()
  qa_report$timestamp <- Sys.time()
  qa_report$data_sources <- data_sources
  
  # 1. Basic data checks
  qa_report$basic_checks <- list(
    total_spatial_units = nrow(results$agricultural),
    missing_classifications = sum(is.na(results$agricultural$N_class))
  )
  
  # 2. Classification distribution
  if ("N_class" %in% names(results$agricultural)) {
    n_dist <- table(results$agricultural$N_class)
    qa_report$classification_distribution <- list(
      nitrogen = n_dist,
      excluded_percentage = round(n_dist["Excluded"] / sum(n_dist) * 100, 1)
    )
  }
  
  # 3. WWTP integration checks
  if ("wwtp" %in% names(results)) {
    qa_report$wwtp_checks <- list()
    
    for (nutrient in names(results$wwtp)) {
      facilities <- results$wwtp[[nutrient]]$facility_data
      qa_report$wwtp_checks[[nutrient]] <- list(
        total_facilities = nrow(facilities),
        facilities_with_loads = sum(!is.na(facilities[[paste0(toupper(substr(nutrient, 1, 1)), "_Load_tons")]]))
      )
    }
  }
  
  # 4. Spatial validation
  if (inherits(results$agricultural, "sf")) {
    qa_report$spatial_checks <- list(
      invalid_geometries = sum(!st_is_valid(results$agricultural)),
      crs = st_crs(results$agricultural)$input
    )
  }
  
  # 5. Range checks
  if ("integrated" %in% names(results)) {
    for (nutrient in names(results$integrated)) {
      data <- results$integrated[[nutrient]]
      surplus_col <- paste0(substr(nutrient, 1, 1), "_surplus")
      
      if (surplus_col %in% names(data)) {
        qa_report$range_checks[[nutrient]] <- list(
          min_surplus = min(data[[surplus_col]], na.rm = TRUE),
          max_surplus = max(data[[surplus_col]], na.rm = TRUE),
          extreme_values = sum(abs(data[[surplus_col]]) > 1000, na.rm = TRUE)
        )
      }
    }
  }
  
  # Generate QA summary
  qa_summary <- list(
    passed = TRUE,
    warnings = character(),
    errors = character()
  )
  
  # Check for issues
  if (qa_report$basic_checks$missing_classifications > 0) {
    qa_summary$warnings <- c(qa_summary$warnings, 
                             paste("Missing classifications:", qa_report$basic_checks$missing_classifications))
  }
  
  if (qa_report$classification_distribution$excluded_percentage > 50) {
    qa_summary$warnings <- c(qa_summary$warnings,
                             paste("High exclusion rate:", qa_report$classification_distribution$excluded_percentage, "%"))
  }
  
  if ("spatial_checks" %in% names(qa_report) && qa_report$spatial_checks$invalid_geometries > 0) {
    qa_summary$errors <- c(qa_summary$errors,
                           paste("Invalid geometries:", qa_report$spatial_checks$invalid_geometries))
    qa_summary$passed <- FALSE
  }
  
  # Print summary
  message("Quality Assurance Summary:")
  message("  Status: ", ifelse(qa_summary$passed, "PASSED", "FAILED"))
  
  if (length(qa_summary$warnings) > 0) {
    message("  Warnings:")
    for (warning in qa_summary$warnings) {
      message("    - ", warning)
    }
  }
  
  if (length(qa_summary$errors) > 0) {
    message("  Errors:")
    for (error in qa_summary$errors) {
      message("    - ", error)
    }
  }
  
  qa_report$summary <- qa_summary
  return(qa_report)
}

# Run QA workflow
# qa_results <- quality_assurance_workflow(results_custom, "Custom 2021 WWTP data")
```

## Getting Help

### Common Questions

**Q: What format should my WWTP data be in?**
A: CSV file with columns for facility name, latitude, longitude, annual load, and state. The package can auto-detect most EPA formats.

**Q: What units can I use for loads?**  
A: "kg", "lbs", "pounds", or "tons". Specify with `wwtp_load_units`.

**Q: Can I use data from other countries?**
A: The spatial boundaries are US-only, but you could provide your own boundaries and modify the coordinate system.

**Q: How do I handle missing data?**
A: The package excludes facilities with missing coordinates or loads. Clean your data first for best results.

**Q: My analysis is running slowly. What can I do?**
A: Try using a smaller spatial scale (HUC2 < HUC8 < County in terms of processing time), fewer years, or clear the data cache with `clear_data_cache()`.

**Q: How do I cite the package and data?**
A: Use `citation_info()` to get proper citations for the package and underlying datasets.

**Q: Can I analyze partial years or seasonal data?**
A: The package is designed for annual data. For seasonal analysis, you would need to aggregate your data to annual totals first.

**Q: What if my WWTP data has different pollutant measurements?**
A: The package expects total nitrogen and total phosphorus loads. If you have other forms (NO3, NH4, PO4), you'll need to convert to total N and P first.

### Function Help

```{r , eval=FALSE}
# Get help on specific functions
?load_user_wwtp
?run_builtin_analysis  
?wwtp_clean_data
?validate_columns

# Check what data is available
check_builtin_data()
list_available_years()

# Package health check
health_check()

# Test data download connection
test_osf_connection()
```

### Getting More Help

- Check the other vignettes: `vignette("getting-started")`, `vignette("visualization-guide")`, `vignette("advanced-features")`
- Look at function documentation: `?function_name`
- Check the package website or GitHub repository for examples
- Use `quick_check()` to validate your results
- Run `health_check()` if you encounter installation or data loading issues

### Reporting Issues

If you encounter bugs or have feature requests:

1. Check that your data format matches the expected format
2. Run `health_check()` to verify package installation
3. Try with built-in data first to isolate the issue
4. Document the error message and steps to reproduce
5. Check the package documentation and vignettes
6. Report issues with a minimal reproducible example

The `manureshed` package is designed to work with your data as easily as possible. Start with the built-in examples, then gradually add your own data as needed. The key is to ensure your data is properly formatted and validated before running the analysis.
